### 2019-08-02 3_implementation.tex
More work on shrinking text.
Update text and fiddle with wording and figures to make everything fit again.

One challenge in implementing Task Bench for such a wide variety of
systems is that the capabilities of the systems vary considerably. For
example, some systems are \emph{implicitly parallel}, and provide some
form of parallelism discovery from sequential programs, whereas others
are \emph{explicitly parallel} and require users to specify the
parallelism in the program. For systems that provide both implicit and explicit parallelism, the form of parallelism used in Task Bench is emphasized in Table~\ref{tab:systems}.

@@

In all cases, members of the respective programming systems' teams
were consulted during the development and evaluation of the
corresponding Task Bench implementations. Where assistance was provided, the insights were invaluable
in helping to ensure that we provide the highest quality implementations for each system.

>>>>

One challenge in targeting such a wide variety of
systems is that the capabilities of the systems vary considerably. For
example, some systems are \emph{implicitly parallel}, and provide some
form of parallelism discovery from sequential programs, whereas others
are \emph{explicitly parallel} and require users to specify the
parallelism in the program. For systems that provide both implicit and explicit parallelism, the form of parallelism used in Task Bench is emphasized in Table~\ref{tab:systems}.

@@

In all cases, members of the programming systems' teams
were consulted in the development and evaluation of the
corresponding Task Bench implementations. Where assistance was provided, the insights helped ensure that we provide the highest quality implementations for each system.

### 2019-08-02 4_metg.tex
Update text and fiddle with wording and figures to make everything fit again.

Existing studies of programming system efficiency
typically use the metric \emph{tasks per second} (TPS). TPS results
are difficult to interpret and apply, because the efficiency (and thus
the amount of useful work) is not constrained. When using empty
tasks~\cite{LegionTracing18}, the resulting upper bound on task
scheduling throughput fails to represent useful work within a
realistic application. With non-empty tasks, since the efficiency of
the overall application is typically not reported~\cite{Canary16,
  Armstrong14}, TPS is not a measurement of runtime-limited
performance. Large tasks may be used to hide any amount of runtime
overhead, while small tasks may result in a drop in total application
throughput even as TPS increases.

@@

To calculate METG, the data is replotted along axes of efficiency
(i.e., as a percentage of the peak FLOP/s achieved) and task
granularity (i.e., $\text{wall time} \times \text{num.~cores}/\text{num.~tasks}$), as shown in Figure~\ref{fig:efficiency-mpi}. Note that a \emph{task} is defined
broadly to be any continuously-executing unit of application code,
and thus it makes sense to discuss tasks even in systems
without an explicit notion of tasking, such as in MPI programs written in a bulk synchronous style. In this case, the tasks run a
compute-bound kernel included in the Task Bench implementation,
described in more detail in Section~\ref{sec:task-bench}.

@@

>>>>

Existing studies of system efficiency
typically report \emph{tasks per second} (TPS). TPS results
are difficult to interpret and apply, because efficiency (and thus
the amount of useful work) is not constrained. With empty
tasks~\cite{LegionTracing18}, the resulting upper bound on task
scheduling throughput fails to represent useful work within a
realistic application. With non-empty tasks, since the efficiency of
the overall application is typically not reported~\cite{Canary16,
  Armstrong14}, TPS is not a measurement of runtime-limited
performance. Large tasks may be used to hide any amount of runtime
overhead, while small tasks may result in a drop in total application
throughput even as TPS increases.

@@

To calculate METG, the data is replotted along axes of efficiency
(i.e., as a percentage of the peak FLOP/s achieved) and task
granularity (i.e., $\text{wall time} \times \text{num.~cores}/\text{num.~tasks}$), as shown in Figure~\ref{fig:efficiency-mpi}. Note that a \emph{task} is defined
broadly to be any continuously-executing unit of application code,
and thus it makes sense to discuss tasks even in systems
with no explicit notion of tasking, such as in MPI programs written in a bulk synchronous style. In this case, the tasks run a
compute-bound kernel included in the Task Bench implementation,
described in more detail in Section~\ref{sec:task-bench}.

@@

\input{e0_flags}

### 2019-08-02 5_evaluation.tex
Update conclusion to match intro.
Update text and fiddle with wording and figures to make everything fit again.

\input{e0_flags}

@@

@@

Figure~\ref{fig:flops} shows the FLOP/s achieved with a compute-bound
kernel with varying problem sizes. This is the full version of
Figure~\ref{fig:flops-mpi}. Each data point in the graph is the mean of 5 runs, with Task Bench configured to execute 1000 time steps of the stencil pattern. In the best case, we measure peak FLOP/s of
$1.26 \times 10^{12}$, which compares favorably with the officially
reported number of $1.2 \times 10^{12}$ \cite{Cori}. For the purposes
of measuring efficiency, we use our empirically determined number as
the baseline for 100\% efficiency.

@@

Figure~\ref{fig:efficiency} plots efficiency (as a percentage of
peak FLOP/s) against task granularity. As described in Section~\ref{sec:metg}, this plot is
used to calculate METG(50\%). The
red, dashed line shows the point where 50\% efficiency is achieved.
In most cases, task granularity asymptotes prior to the 50\% efficiency point,
though some systems continue to improve as efficiency continues
to drop. Accounting for this effect is one of the main arguments in
favor of METG with a reasonable efficiency threshold instead of
measuring task scheduling throughput of empty tasks
(effectively METG(0\%)). Measuring performance using empty tasks can
reward implementation strategies, such as devoting nearly 100\% of
system resources to the runtime system, that make no sense for real
applications.

@@

Figure~\ref{fig:bytes} shows performance with a memory-bound kernel. We measure peak memory
bandwidth of 108~GB/s, which compares favorably with the
OpenMP-enabled STREAM benchmarks~\cite{STREAM} which report up to 98~GB/s.

Not all cores are required to saturate memory bandwidth. As a result,
systems that reserve one or more cores are able to hit 100\% of peak
performance, unlike with compute-bound kernels. We are currently investigating the
performance of Charm++, Regent and X10.

% FIXME: why are Charm++, Regent and X10 slow?

@@

The ratio in METG between 0 and 3 dependencies per task ranges from
$1.01\times$ to $250\times$ (median $2.9\times$). The difference is most pronounced in systems that
perform runtime work inline. For example, MPI achieves an METG of 390
ns with 0 dependencies, but this rises to 4.6 \textmu{}s with 3 dependencies,
a $12\times$ increase. This is unsurprising as in the case
of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
issued at all, so MPI has nothing to do aside from executing kernels
as quickly as possible. For this reason, we recommend evaluating the performance of
programming systems under realistic workloads, rather than with
trivial dependencies.

>>>>

@@

Figure~\ref{fig:flops} shows the FLOP/s achieved with a compute-bound
kernel with varying problem sizes (simulated by running the kernel for varying numbers of iterations). This is the full version of
Figure~\ref{fig:flops-mpi}. Each data point in the graph is the mean of 5 runs, with Task Bench configured to execute 1000 time steps of the stencil pattern. In the best case, we measure peak FLOP/s of
$1.26 \times 10^{12}$, which compares favorably with the officially
reported number of $1.2 \times 10^{12}$ \cite{Cori}. For the purposes
of measuring efficiency, we use our empirically determined number as
the baseline for 100\% efficiency.

@@

@@

Figure~\ref{fig:efficiency} plots efficiency (as a percentage of
peak FLOP/s) against task granularity. As described in Section~\ref{sec:metg}, this plot is
used to calculate METG(50\%). The
red, dashed line shows the point where 50\% efficiency is achieved.
In most cases, task granularity asymptotes prior to the 50\% efficiency point,
though some systems continue to improve as efficiency drops further. Accounting for this effect is one of the main arguments in
favor of METG with a reasonable efficiency threshold instead of
measuring task scheduling throughput of empty tasks
(effectively METG(0\%)). Measuring performance using empty tasks can
reward implementation strategies, such as devoting nearly 100\% of
system resources to the runtime system, that make no sense for real
applications.

@@

Figure~\ref{fig:bytes} shows performance with a memory-bound kernel. We measure a peak memory
bandwidth of 79~GB/s, using a working set size of 0.5 GB. As discussed in Section~\ref{sec:task-bench}, the kernels are designed to keep the working set constant as the number of iterations decrease to avoid noisy, superlinear effects in the results. For comparison, the
OpenMP-enabled STREAM benchmarks~\cite{STREAM} report up to 98~GB/s.

Not all cores are required to saturate memory bandwidth, reducing the
impact of reserving cores for internal use. Most systems hit 100\% of
peak, unlike the compute-bound case.

@@

The ratio in METG between 0 and 3 dependencies per task ranges from
$1.01\times$ to $250\times$ (median $2.9\times$). The difference is most pronounced in systems that
perform runtime work inline. For example, MPI achieves an METG of 390
ns with 0 dependencies, but this rises to 4.6 \textmu{}s with 3 dependencies,
a $12\times$ increase. This is unsurprising as in the case
of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
issued at all, so MPI has nothing to do aside from executing kernels
as quickly as possible. Clearly, choosing a representative dependence
pattern is important when estimating the performance of a workload or
class of workloads.
